---
title: "Running a simple simulation"
description: "Get Started with MSTWeatherGen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Running a simple simulation}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
editor_options:
    chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```


## Introduction

This vignette provides a step-by-step guide on running a stochastic weather generator using the `MSTWeatherGen` package. From loading the historical weather data and spatial coordinates, performing parameter estimation, to running simulations and generating validation plots. This guide covers all you need to get started with `MSTWeatherGen`, but does not provide the technical details of the methods considered.

## Load the `MSTWeatherGen` Package

First, let's load the `MSTWeatherGen` package.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(MSTWeatherGen)
```

## Data

### Toy data

For our simulation, we will need historical weather data of multiple variables and geographic coordinates. Below, we load these datasets stored within the package.

The data considered here is the meteorological dataset SAFRAN, developed by Météo-France. We only consider a small region in the south of France for the period 2017-2021. We consider 3 variables: precipitation, wind, and maximum temperature.

```{r echo=TRUE}
data("data", package = "MSTWeatherGen")
data("coordinates", package = "MSTWeatherGen")
#load("/home/sobakrim/Bureau/MSTWeatherGen/MSTWeatherGen/data/real_data.rda")
#load("/home/sobakrim/Bureau/MSTWeatherGen/MSTWeatherGen/data/coordinates.rda")
names = c("Precipitation","Wind","Temp_max")
dates = seq(as.Date("2017-01-01"),as.Date("2021-12-31"), by="day")
```

### Note on data needed for `MSTWeatherGen`

`MSTWeatherGen` is designed for multivariate and spatio-temporal weather generation. Therefore, the meteorological data to be used (here `data`) needs to be multiple variables (at least two variables) defined in a spatial domain (characterized by `coordinates`) and a temporal domain (defined by `dates`). Thus, the data needs to be in three dimensions (time$\times$location$\times$variable). Alongside the meteorological data, the user needs to provide a two column matrix of coordinates of each locations considered. Each row `i` of this matrix (`coordinates[i,]`) corresponds to the x and y coordinates of the datum `data[,i,]`. Finally, the user needs also to provide the time (`dates` here) corresponding to each point in the first dimension of `data`.

Another important point about the meteorological data is the precipitation variable. Being a special variable in the estimation and simulation (as it has many zeros), if it is considered, it has to be the first variable in the meteorological data. If the user wants, one can also provide the names of the variables considered (here `names`).

## Estimation

With the data loaded, we proceed to estimate the parameters required for simulation. This involves determining weather types, computing transition probabilities between weather types, estimating the transformation functions of the variables into normal distribution, and finally estimating the parameters of the multivariate spatio-temporal covariance function, all using the `MSTWeatherGen.Estim` function.

Before heading to the estimation, we will detail some important considerations that will help the user use the `MSTWeatherGen.Estim` function. First, if the data exhibit seasonality, it has to be handled somehow. Here, propose to treat each season separately meaning that the parameters of the model (weather types, transformation functions, and covariance function) are estimated at each season. The seasons are up to the user to define using the following format:

```{r echo=TRUE}
seasons <- list(
  s1 = list(min_day = 1, max_day = 29, min_month = 12, max_month = 2),
  s2 = list(min_day = 1, max_day = 31, min_month = 3, max_month = 5),
  s3 = list(min_day = 1, max_day = 31, min_month = 6, max_month = 8),
  s4 = list(min_day = 1, max_day = 30, min_month = 9, max_month = 11)
)
```

We can also provide the names of each season, to be used later for the validation plots:

```{r echo=TRUE}
names_seasons = c("DJF", "MAM", "JJA", "SON")
```

If the considered data does not exhibit seasonality one can use the whole period to estimate the parameters. To do that, one needs to fix the argument `by_season` of the `MSTWeatherGen.Estim` as `False`.

All the data being ready, we can now proceed to the estimation of the parameters using `MSTWeatherGen.Estim` function.

```{r echo=TRUE, message=FALSE, warning=FALSE}
names_weather_types = c("Wind", "Temp_max")
swg = MSTWeatherGen.Estim(data = data, seasons = seasons,dates = dates, names = names, by_season = T, 
                        precipitation = T,names_weather_types = names_weather_types,
                        coordinates= coordinates, max_it=100, tmax=2, n1=5, n2=5)
```

Along the important arguments already defined (`data`, `seasons`, `by_season`, `dates`, `names`, `coordinates`), the user needs to define other arguments related to the estimation such as `max_it` which is the maximum iteration used by the `optim` function for the estimation of the covariance parameters, `tmax` which is the maximum time lag considered in the covariance estimation, and `n1` and `n2` which are related to the selection of the spatial pairs in the covariance estimation. Finally, the user needs to specify if the precipitation variable is considered using the `precipitation` argument. The `names_weather_types` is optional and it provides the names of the variables to be used for the weather types determination. If not provided, all the variables are used.

Now that the estimation is done, which may take some time depending on the size of the data, we can use the estimated parameters (stored in `swg` object) for the simulation.

## Simulation

Now, we're ready to simulate weather data based on the estimated parameters. The simulation method used by `MSTWeatherGen` is based on the vector autoregressive (AR) model whose coefficients matrices can be calculated using the covariance parameters estimated using `MSTWeatherGen.Estim` function. Therefore, before running the simulation, the user needs to define the lag of the AR model `AR_lag` and then calculate the coefficient matrices using `calculate_AR_coefficients_matrices`. Note that choosing a higher value of `AR_lag` can be computationally challenging both in term of calculation and storage. Therefore, the choice of this parameters depends of the size of the spatial domain and the number of variables considered. Here we choose an AR of order 2 and fix `AR_lag` as 2.

```{r echo=TRUE}
AR_lag = 2
bk = calculate_AR_coefficients_matrices(parm = swg, coordinates = coordinates, AR_lag = AR_lag)
```

After calculating the AR coefficient matrices, we can now ran the simulation using `MSTWeatherGen_Sim` function.

```{r echo=TRUE}
sim = MSTWeatherGen_Sim(dates_sim = dates, dates_original = dates, 
                                data = data, seasons = seasons, parm = swg, AR_lag = AR_lag, bk = bk)
```

The object `sim` contains the simulated data in a 3D (time$\times$location$\times$variable) format (such as `data`) for the `dates_sim` dates.

## Validation plots

Finally, to validate our simulation, we can compare the simulated data against our historical data. Here are some plots that help in assessing the accuracy and quality of the simulation.

We first choose a location to use for validation.

```{r echo=TRUE}
Avignon <- which.min(apply(coordinates, 1, function(point) {
  sqrt(sum((point - c(4.8055, 43.9493))^2))
}))
```

We now plot the observed empirical density versus simulated weather variables in Avignon during the seasons considered (winter (DJF), spring (MAM), summer (JJA), and fall (SON)) using `plot_observed_vs_simulated_density`function

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
plot_observed_vs_simulated_density(sim = sim, observed = data, dates = dates, 
                              seasons = seasons, location = Avignon, names = names, names_seasons =names_seasons)

```

Next, we can compare the observed and simulated monthly variable means in Avignon using `plot_mean_by_month` function

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
plot_mean_by_month(sim = sim, observed = data,places = Avignon, names_places = "Avignon", names = names, dates = dates)
```

Then, we can compare simulated and observed frequency of wet days in the considered area during the considered seasons

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
plot_wet_frequency(sim, data, dates, seasons, coordinates, names_seasons)
```

Finally, we can see if our stochastic weather generator can reproduce the observed length of wet spells in the considered region using the `plot_dry_wet_spells_maps` function

```{r echo=TRUE, message=FALSE, warning=FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
id = lubridate::year(dates) %in% 2018:2021
plot_dry_wet_spells_maps(sim = sim[id,,], observed = data[id,,], coordinates = coordinates, dates = dates[id])
```
